---
title: "EDA Maternal Healthcare Disparities"
output: html_document
date: "2024-06-07"
---

```{r setup, include=FALSE}

```

## Overview of the Dataset 
The dataset that we were given was maternal healthcare disparities. It is
a sample of data from the CDC database for birth records from 2019. It includes details such as the number of prior births now deceased, maternal tobacco use, and whether the mother had pre-pregnancy diabetes or hypertension. This dataset provides statistics for births, including the average age of the mother, average birth weight, average pre-pregnancy BMI, average number of prenatal visits, and the average interval since the last birth. This data allows for analysis of maternal and infant health trends across different states. It offers us valuable insights into the impacts of these factors on birth outcomes.


```{r}
library(tidyverse)
maternal <- read_csv("https://raw.githubusercontent.com/36-SURE/36-SURE.github.io/main/data/maternal.csv")

```

## Question 1: 
Is higher maternal age associated with higher prenatal visits? 

```{r}
##Question 2:
Is there a relationship between tobacco use and pre-pregnancy conditions like diabetes and hypertension?

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
