---
title: "EDA Maternal Healthcare Disparities"
output: html_document
date: "2024-06-07"
---

```{r setup, include=FALSE}

```

## Overview of the Dataset 
The dataset that we were given was maternal healthcare disparities. It is
a sample of data from the CDC database for birth records from 2019. It includes details such as the number of prior births now deceased, maternal tobacco use, and whether the mother had pre-pregnancy diabetes or hypertension. This dataset provides statistics for births, including the average age of the mother, average birth weight, average pre-pregnancy BMI, average number of prenatal visits, and the average interval since the last birth. This data allows for analysis of maternal and infant health trends across different states. It offers us valuable insights into the impacts of these factors on birth outcomes.


```{r}
library(tidyverse)
maternal <- read_csv("https://raw.githubusercontent.com/36-SURE/36-SURE.github.io/main/data/maternal.csv")

```

## Question 1: 
Is higher maternal age associated with higher prenatal visits? To look at the relationship between these two variables, we can use a scatterplot. 

```{r}
maternal %>%  
  ggplot(aes(x=AverageMotherAge, y=AverageNumberPrenatalVisits)) +
  geom_point(color="lightpink",size=3,alpha=0.5) +
  theme_light() 
# adding a linear regression line 
maternal %>% 
  ggplot(aes(x = AverageMotherAge, y = AverageNumberPrenatalVisits)) +
  geom_point(color = "lightpink", size = 3, alpha = 0.5) +
  geom_smooth(method = "lm", linewidth = 2, color="hotpink") +
  theme_light
```

The graphs show that there seems to be some positive relationship between the Average Mother Age and Average Prenatal Visits. More convincing proof would be to check the correlation coefficient. 

```{r}
cor(maternal$AverageMotherAge, 
    maternal$AverageNumberPrenatalVisits, 
    use = "complete.obs")
```
Now looking at the same data but with clustering with kmeans to see any other significant findings. 

```{r}
clean_data <- maternal %>%  
  select(AverageMotherAge, AverageNumberPrenatalVisits) %>%
  na.omit()

init_kmeans <- clean_data %>%  
  kmeans(centers = 5, nstart = 1, algorithm = "Lloyd")

clean_data <- clean_data %>%
  mutate(clusters = as.factor(init_kmeans$cluster)) #yo

clean_data %>%
  ggplot(aes(x = AverageMotherAge, y = AverageNumberPrenatalVisits, color = clusters)) +
  geom_point() + 
  labs(title = "Clustering" ,
       x = "Average Mother Age",
       y = "Average Number of Prenatal Visits") +
  coord_fixed()

ggplot(clean_data, aes(x = clusters, y = AverageNumberPrenatalVisits, color = clusters)) +
  geom_boxplot() +
  labs(x = "Cluster", y = "Average Prenantal Visits") +
  theme_light()

ggplot(clean_data, aes(x = clusters, y = AverageMotherAge, color = clusters)) +
  geom_boxplot() +
  labs(x = "Cluster", y = "Average Mother Age") +
  theme_light()

```

## Question 2: 
Is there an effect of tobacco use and pre-existing diabetes in the average birth weight of a newborn?

```{r}
library(tidyverse)

maternal <- read_csv("https://raw.githubusercontent.com/36-SURE/36-SURE.github.io/main/data/maternal.csv")


subset_data <- maternal %>%
  select(TobaccoUse, PrePregnancyDiabetes, AverageBirthWeight)


subset_data <- subset_data %>%
  mutate(
    TobaccoUse = factor(TobaccoUse),
    PrePregnancyDiabetes = factor(PrePregnancyDiabetes)
  )

ggplot(subset_data, aes(x = TobaccoUse, y = AverageBirthWeight, fill = PrePregnancyDiabetes)) +
  geom_boxplot() +
  labs(
    title = "Effect of Tobacco Use and Pre-Pregnancy Diabetes on Newborns' Birth Weight",
    x = "Tobacco Use",
    y = "Average Birth Weight",
    fill = "Pre-Pregnancy Diabetes"
  ) +
  scale_fill_manual(values = c("No" = "lightblue", "Yes" = "Hotpink", "Unknown" = "lightgrey")) +
  theme_light()
```

Now with clustering 

```{r}


```

## Question 3: How does the number of births in each state with Tobacco Use, Pre Pregnancy Diabetes, and Pre Pregnancy Hypertension differ with the marginal difference of not using tobacco.
```{r}
library(tidyverse)
library(ggplot2)
maternal <- read_csv("https://raw.githubusercontent.com/36-SURE/36-SURE.github.io/main/data/maternal.csv")

#View(maternal)

#How does the number of births in each state with Tobacco Use, Pre Pregnancy Diabetes, and Pre Pregnancy Hypertension differ with the marginal difference of not using tobacco.


library(ggthemes)
#1: Maternal tobacco use is associated with lower average birthweights
#2: Does maternal age correlate with birthweight?
#3: How do number of births in that state with a defined combination of the previous four conditions (PriorBirthsNowDeceased, TobaccoUse, PrePregnancyDiabetes, PrePregnancyHypertension) vary between each state

# maternal |> 
#   filter(TobaccoUse %in% c("Yes"), PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("Yes")) |> 
#   select(c("State","Births")) |> 
#   mutate(State=factor(State),
#          State = fct_reorder(State, Births)) |> 
#   ggplot() +
#   geom_segment( aes(x=State, xend=State, y=0, yend=Births)) +
#   coord_flip()+
#   #theme_fivethirtyeight() +
#   theme(
#     panel.grid.major.x = element_blank(),
#     panel.border = element_blank(),
#     axis.ticks.x = element_blank(),
#     axis.text.x =  
#   ) +
#   geom_point(aes(x=State,y=Births),size=3, color="red", fill=alpha("orange", 0.3), alpha=0.6, shape=21, stroke=2)+
#   labs(
#      x = "State",
#      y = "Number of births in each state with TobaccoUse, PrePregnancyDiabetes, and PrePregnancyHypertension" 
#   )

tobacco_no = maternal |> 
  filter(TobaccoUse %in% c("No"), PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("Yes"))

tobacco_yes = maternal |> 
  filter(TobaccoUse %in% c("Yes"), 
         PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("Yes"))

tobacco_equals = rbind(tobacco_no, tobacco_yes)
tobacco_equals |> 
  select(c("State","Births","TobaccoUse")) |> 
  mutate(State=factor(State),
         State = fct_reorder(State, Births)) |> 
  ggplot(aes(group=State)) +
  geom_segment( aes(x=State, xend=State, y=0, yend=Births)) +
  geom_point(aes(x=State,y=Births, fill = TobaccoUse),size=5, 
             alpha=0.67, shape=21)+
  scale_y_continuous(breaks=seq(0,600,50))+
  coord_flip()+
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(size = 13, hjust = 0, vjust = 0.5, face = "bold", 
                              margin = margin(b = 0.2, unit = "cm"))
  ) +
  labs(
    title = "Number of Births in each State: Tobacco Use vs. Non-Use, Considering Diabetes and Hypertension",
    x = "State",
    y = "Births",
    fill = "Tobacco Use"
  )

 
#first filter then add another filter 
#rbind f



#Diabetes
diabetes_no = maternal |> 
  filter(TobaccoUse %in% c("Yes"), PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("No"), PrePregnancyHypertension %in% c("Yes"))
diabetes_yes = maternal |> 
  filter(TobaccoUse %in% c("Yes"), 
         PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("Yes"))

diabetes_equals = rbind(diabetes_no,diabetes_yes)
diabetes <- diabetes_equals |> 
  select(c("State","Births","PrePregnancyDiabetes")) |> 
  mutate(State=factor(State),
         State = fct_reorder(State, Births)) |> 
  ggplot(aes(group=State)) +
  geom_segment( aes(x=State, xend=State, y=0, yend=Births)) +
  geom_point(aes(x=State,y=Births, fill = PrePregnancyDiabetes),size=5, 
             alpha=0.67, shape=21)+
  scale_y_continuous(breaks=seq(0,600,50))+
  coord_flip()+
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(size = 13, hjust = 0, vjust = 0.5, face = "bold", 
                             margin = margin(b = 0.2, unit = "cm"))
  ) +
  labs(
    title = "Number of Births in each State: With or Without Diabetes, Considering Tobacco Use and Hypertension",
    x = "State",
    y = "Births",
    fill = "Diabetes"
  )

diabetes


#Hypertension

hypertension_no = maternal |> 
  filter(TobaccoUse %in% c("Yes"), PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("No"))
hypertension_yes = maternal |> 
  filter(TobaccoUse %in% c("Yes"), 
         PriorBirthsNowDeceased == '0', 
         PrePregnancyDiabetes %in% c("Yes"), PrePregnancyHypertension %in% c("Yes"))

hypertension_equals = rbind(hypertension_no,hypertension_yes)
hypertension <- hypertension_equals |> 
  select(c("State","Births","PrePregnancyHypertension")) |> 
  mutate(State=factor(State),
         State = fct_reorder(State, Births)) |> 
  ggplot(aes(group=State)) +
  geom_segment( aes(x=State, xend=State, y=0, yend=Births)) +
  geom_point(aes(x=State,y=Births, fill = PrePregnancyHypertension),size=5, 
             alpha=0.67, shape=21)+
  scale_y_continuous(breaks=seq(0,600,50))+
  coord_flip()+
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(size = 13, hjust = 0, vjust = 0.5, face = "bold", 
                              margin = margin(b = 0.2, unit = "cm"))
  ) +
  labs(
    title = "Number of Births in each State: With or Without Hypertension, Considering Tobacco Use and Pre-Pregnancy Diabetes",
    x = "State",
    y = "Births",
    fill = "Diabetes"
  )
hypertension


```

=======
## Question 3: 
>>>>>>> ba032a28603a90bf956817e083fc07d66e999972

